# WAT PROCESSING
This script will generate the following figures:
Fig. EV1A- WAT UMAP without non-adipocells, grouped by original clusterings
Fig. EV1B- WAT UMAP without non-adipocells before CC regression, grouped by CC Phase
Fig. EV1C- WAT cell cycle distribution pre-regression
Fig. EV1D - WAT UMAP without non-adipocells after CC regression, grouped by CC Phase
Fig. EV1E - WAT cell cycle distribution post-regression
Fig. EV2A - markers of top 10% wAPC heatmap
Fig. 1B Pre-exclusion UMAP of WAT
Fig. 1D Plot violin plots of adipocells and non-adipocell markers 
Fig. 1F - Final UMAP of WAT clusters after non-adipocell exclusion and CC regression 
Fig. 1H - Stacked violin plots of prominent markers
Fig. 2D Feature plots 
Fig. 2F Regulon score patterns
Table EV1 - wAPC markers
Table EV3 - DEGs between wCAPs vs wFAPLs

#A. Seurat processing

## 1. Load in libraries
```{r}
library(Seurat)
library(readr)
library(dplyr)
library(patchwork)
library(data.table)
library(ggplot2)
library(cowplot)
library(pheatmap)
library(sctransform)
library(SeuratDisk)
library(scater)
library(tidyverse)
library(SeuratWrappers)
library(gplots)
library(stringr)
library(scCustomize)
```

## 2. Load in data
```{r}
wat.data = Read10X(data.dir = "//research.drive.wisc.edu/galmozzi/Galmozzi lab/Hoang/Projects/Adipocyte Subtypes/new_analysis/0WAT/filtered_feature_bc_matrix")

```

```{r}
#convert into seurat object
wat = CreateSeuratObject(counts = wat.data, project = "0watpreads", min.cells = 3, min.features = 200)
wat$data.set = rep("wat", length(wat$orig.ident))
```

## 3. Calculate % mitochondria RNA and apply filters
```{r}
#calculate percent mitochondrial RNA for watpreads
wat[["percent.mt"]] = PercentageFeatureSet(wat, pattern = "^mt-")

#generate violin plot #genes/cell, mtRNA/cell, reads/cell
VlnPlot(wat, features= c("nFeature_RNA", "nCount_RNA", "percent.mt"), pt.size=0.5, ncol = 3)


#apply filter-- UMI 6000-60000, features 3900-7700, mito umi 1.7-5
wat <- subset(wat, subset = nFeature_RNA > 3900 & nFeature_RNA < 7700 & percent.mt < 5& percent.mt > 1.7 & nCount_RNA > 6000 & nCount_RNA < 60000)
```

## 4. Data normalization and scaling
```{r}
#normalizing the data -normalizes the feature expression measurements for each cell by the total expression, multiplies this by a scale factor (10,000 by default), and log-transforms the result
wat <- NormalizeData(wat, normalization.method = "LogNormalize", scale.factor = 10000)

#identification of highly variable features
wat <- FindVariableFeatures(wat, selection.method = "vst", nfeatures = 2000)

#scaling the data 
all.genes <- rownames(wat)
wat <- ScaleData(wat, features = all.genes)
```

## 5. Run PCA and UMAP clustering
```{r}
#PCA
wat <- RunPCA(wat, features = VariableFeatures(object = wat))
#clustering
wat <- FindNeighbors(wat, dims = 1:10)
wat <- FindClusters(wat, resolution = 0.5)
#umap
wat <- RunUMAP(wat, dims = 1:10)
DimPlot(wat, reduction = "umap")
```

## 6. Find all markers to identify non-adipocells

```{r}
#find all markers
wat.markers<- FindAllMarkers(wat) 

#renaming identities for subclusters based on the markers identified
new_cluster_ids <- c('APCs 1', 'APCs 2', "APCs 3", "APCs 4",'APCs 5', 'APCs 6', 'Glial cells' , "Myeloid cells")
names(new_cluster_ids) <- levels(wat)
wat <- RenameIdents(wat, new_cluster_ids)

#create a new metadata column called "wat_unexcluded_clusters"
wat@meta.data$wat_unexcluded_clusters <- as.factor(wat@active.ident) 
```

Save clustering information
```{r}
write.csv(wat$wat_unexcluded_clusters, "F:/new_analysis/scripts/publication/wat_unexcluded_clusters.csv")
write.csv(wat@reductions$pca@cell.embeddings, "F:/new_analysis/scripts/publication/wat_unexcluded_pca.csv")
write.csv(wat@reductions$pca@feature.loadings, "F:/new_analysis/scripts/publication/wat_unexcluded_pca_loadings.csv")
write.csv(wat@reductions$umap@cell.embeddings, "F:/new_analysis/scripts/publication/wat_unexcluded_umap.csv")
```


## 7. Remove non-adipocells and recluster
```{r}
#redefine the wat object as wat_unexclude for future reference
wat_unexcluded <- wat

wat_excluded <- subset(x=wat_unexcluded, 
                      idents = c("Glial cells", "Myeloid cells"),
                      invert = TRUE) #remove glial and myeloid cells

wat_excluded <- NormalizeData(wat_excluded, normalization.method = "LogNormalize", scale.factor = 10000)
wat_excluded <- FindVariableFeatures(wat_excluded, selection.method = "vst")
wat_excluded <- RunPCA(wat_excluded, features = VariableFeatures(object = wat_excluded))
wat_excluded <- FindNeighbors(wat_excluded, dims = 1:10)
wat_excluded <- FindClusters(wat_excluded, resolution = 0.5)
wat_excluded <- RunUMAP(wat_excluded, dims = 1:10)
```


## 8. Cell cycle regression
## Cell cycle scoring
```{r}
#rename wat_excluded to wat_nocc to track cell cycle status
wat_nocc = wat_excluded

#cell cycle scoring
m.s.genes <- str_to_title(cc.genes$s.genes)
m.g2m.genes <- str_to_title(cc.genes$g2m.genes)
wat_nocc <- CellCycleScoring(wat_nocc, s.features = m.s.genes, g2m.features = m.g2m.genes, set.ident = FALSE)
```
## Cell cycle regression
```{r}
#REGRESS OUT CELL CYCLE SCORES DURING DATA SCALING
wat_cc <- ScaleData(wat_nocc, vars.to.regress = c("S.Score", "G2M.Score"), features = rownames(wat_nocc))
#run PCA 
wat_cc <- RunPCA(wat_cc)
#RUN UMAP ON CC 
wat_cc <- FindNeighbors(wat_cc, dims = 1:10)
wat_cc<- FindClusters(wat_cc, resolution = 0.5)
wat_cc <- RunUMAP(wat_cc, dims = 1:10)
```


## 9. final cluster names
```{r}
Idents(wat_cc) <- "seurat_clusters"
new_cluster_ids <- c('wCAPs', 'Early Preads W1', "Early Preads W2", "Early Preads W3",'wFAPLs', 'Early Preads W4', 'Differentiating wFAPLs')
names(new_cluster_ids) <- levels(wat_cc)
wat_cc <- RenameIdents(wat_cc, new_cluster_ids)
wat_cc@meta.data$wat_named_wclusters <- as.factor(wat_cc@active.ident)



wat_cc$wat_named_wclusters <- factor(wat_cc$wat_named_wclusters, levels = c('Early Preads W1', "Early Preads W2", "Early Preads W3", "Early Preads W4","wCAPs",'wFAPLs','Differentiating wFAPLs'))
Idents(wat_cc) <- "wat_named_wclusters"
```



## 10. Find markers of each cluster
```{r}
wat_cc <- SCTransform(wat_cc)
wcaps_markers= FindMarkers(wat_cc,'wCAPs')
wfapls_markers= FindMarkers(wat_cc, 'wFAPLs')
dwfapls_markers= FindMarkers(wat_cc, 'Differentiating wFAPLs')
w1_markers = FindMarkers(wat_cc, 'Early Preads W1')
w2_markers= FindMarkers(wat_cc, 'Early Preads W2')
w3_markers= FindMarkers(wat_cc, 'Early Preads W3')
w4_markers= FindMarkers(wat_cc, 'Early Preads W4')
```

Filter for top 10% markers
```{r}
w1_markers_filtered <- w1_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.75)  %>%
  arrange(desc(avg_log2FC)) %>%  # Sort wy avg_log2FC in descending order
  slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC       

w2_markers_filtered <- w2_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.75) %>%
  arrange(desc(avg_log2FC)) %>%  
  slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC 

w3_markers_filtered <- w3_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.75) %>%
  arrange(desc(avg_log2FC)) %>%  
  slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC 

w4_markers_filtered <- w4_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.75) %>%
  arrange(desc(avg_log2FC)) %>%  
    slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC   
dwfapls_markers_filtered <- dwfapls_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.75) %>%
  arrange(desc(avg_log2FC)) %>%  
    slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC   
wcaps_markers_filtered <- wcaps_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.75) %>%
  arrange(desc(avg_log2FC)) %>%  
    slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC   
wfapls_markers_filtered <- wfapls_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.75) %>%
  arrange(desc(avg_log2FC)) %>%  
    slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC 


#extract just the name of genes

dwfapls_markers_list <-rownames(dwfapls_markers_filtered)
wcaps_markers_list <- rownames(wcaps_markers_filtered)
wfapls_markers_list <- rownames(wfapls_markers_filtered)
w1_markers_list <-rownames(w1_markers_filtered)
w2_markers_list <- rownames(w2_markers_filtered)
w3_markers_list <- rownames(w3_markers_filtered)
w4_markers_list <- rownames(w4_markers_filtered)

#concatenate all significant markers into 1 single vector
significant_markers_wat= c(w1_markers_list,w2_markers_list,w3_markers_list,w4_markers_list, wcaps_markers_list,wfapls_markers_list,dwfapls_markers_list)
```

## 11. Export UMAP and metadata for RNA velocity analysis
```{r}
#export uMAP projections
umap = cbind("Barcode" = rownames(Embeddings(object = wat_cc, reduction = "umap")), Embeddings(object = wat_cc, reduction = "umap"))
write.table(umap, file="F:/uMAP_seurat.csv", sep = ",", quote = F, row.names = F, col.names = T)

#export metadata to get clusters export
write.csv(wat_cc@meta.data,"F:/seurat_metadata.csv") #use excel to extract barcodes and cluster names later

```



# B. Scenic analysis
Here, we will be using the aucell score scenic output for WAT, merging the matrix with the wat_cc object and perform DE expression for AUCell differential scores.

## 1. Load in libraries
library(Seurat)
library(dplyr)


## 2. Load in data
```{r}
#wat <- readRDS('F:/new_analysis/labeled_analysis/wat.rds')
aucell_scores <- read.csv("F:/scenic/wat/data/auc_mtx_v10.csv", row.names = 1)

seurat_object=wat_cc #set current object to be wat
```

## 3. Process the AUCell matrix
```{r}
# Cells should be columns, regulons should be rows
if (!identical(colnames(aucell_scores), colnames(seurat_object))) {
  aucell_scores <- t(aucell_scores)
}
# Set proper names if needed
# Create a new assay object with AUCell scores
aucell_assay <- CreateAssayObject(counts = aucell_scores)

# Add this new assay to your Seurat object
seurat_object[["AUCell"]] <- aucell_assay
DefaultAssay(seurat_object) <- "AUCell"
seurat_object<- ScaleData(seurat_object)
```

## 4. Find differential scores
```{r}
wcaps_tcf <- FindMarkers(seurat_object, ident.1='wCAPs', ident.2='wFAPLs') 
wcaps_tcf_filtered <- wcaps_tcf %>%
  filter(p_val_adj < 0.05 & avg_log2FC>0.1 & pct.1>0.5)

wcaps_tcf_list <- rownames(wcaps_tcf_filtered)

wfapls_tcf <-FindMarkers(seurat_object, ident.1='wFAPLs', ident.2='wCAPs')
wfapls_tcf_filtered <- wfapls_tcf %>%
  filter(p_val_adj < 0.05 & avg_log2FC>0 & pct.1>0.5)
wfapls_tcf_list <- rownames(wfapls_tcf_filtered)
```

## 5. Filter out TCF that shows erratic expressions
```{r}
#FINAL HEATMAPS 3.5.25
wcaps_tcf_to_remove <- c('Osr2...', 'Satb1...', 'Tbx1...', 'Zfp282...','Nfatc2...','Dlx5...','Hoxa3...', 'Nr1d2...','Klf10...') #adipogenic
wfapls_tcf_to_remove <- c('Deaf1...','Irx2...', 'Irf1...','E2f1...','Tfdp1...', 'Sirt6...', 'Hoxb8...', 'Hoxc8...','Hoxa9...','Hoxc9...','Hoxb2...','Xbp1...','Erf...','Sox18...','Pura...','Spi1...',' Ikzf1...', 'Zfhx3...' ) #Fibrogenic

# Remove the genes from the list
adipogenic_tcf= wcaps_tcf_list
adipogenic_tcf_cleaned <- adipogenic_tcf[!adipogenic_tcf %in% wcaps_tcf_to_remove] 


fibrogenic_tcf= wfapls_tcf_list
fibrogenic_tcf_cleaned <- fibrogenic_tcf[!fibrogenic_tcf %in% wfapls_tcf_to_remove] 

```

## 6. subset data for heatmap plotting
```{r}
#extract the data matrix from seurat object
data_matrix <- GetAssayData(object = seurat_object, assay = "AUCell", slot='scale.data') 
#setting up the labels and colors
cluster_labels <- seurat_object$wat_named_wclusters
# Order data matrix by cluster labels
ordered_indices <- order(cluster_labels)
ordered_data_matrix <- data_matrix[, ordered_indices]
ordered_cluster_labels <- cluster_labels[ordered_indices]
# Create annotation data frame
cluster_annotation <- data.frame(Cluster = factor(ordered_cluster_labels))

#define color for each cluster
cluster_colors<- setNames(c('#00cd66','#fcd589',"#cd6600","#ff3030", '#8c510a','#5ab4ac','#01665e'), sort(unique(cluster_labels)))
# Create the annotation data frame
cluster_annotation <- data.frame(Cluster = factor(ordered_cluster_labels, levels = names(cluster_colors)))


#Set limits and breaks based on standard deviation
# Calculate the standard deviation of the matrix
std_dev <- sd(as.vector(data_matrix))
# Define breaks around zero, within ±2 standard deviations
max_limit <- 2 * std_dev
breaks <- seq(-max_limit, max_limit, length.out = 101)
```

Select CAPs or FAPs heatmap
```{r}
regulon_activity_caps <- ordered_data_matrix[adipogenic_tcf_cleaned, ] #for CAPs heatmap
regulon_activity_fapls <- ordered_data_matrix[fibrogenic_tcf_cleaned, ] # for FAPLs heatmap

```

Plot the heatmap
```{r}
pheatmap(regulon_activity_caps, #or regulon_activity_fapls
         cluster_cols = TRUE,show_colnames = FALSE,
         cluster_rows = TRUE,
         scale = 'none',  # No scaling, to keep data as is
         breaks=breaks,
         annotation_col = cluster_annotation,
         annotation_colors = list(Cluster = cluster_colors))
```


# FIGURE PLOTS

Fig. EV1A- WAT UMAP without non-adipocells, grouped by original clusterings
```{r}
DimPlot(wat_nocc, reduction = "umap", cols=wat_raw_colors, group.by = "wat_unexcluded_clusters", pt.size=2)
```

Fig. EV1B- WAT UMAP without non-adipocells before CC regression, grouped by CC Phase
```{r}
DimPlot(wat_nocc, reduction = "umap", group.by = "Phase", pt.size=2)
```

Fig. EV1C- WAT cell cycle distribution pre-regression
```{r}
cell_counts <- table(wat_nocc$wat_unexcluded_clusters, wat_nocc$Phase)
print(cell_counts)

```

Fig. EV1D - WAT UMAP without non-adipocells after CC regression, grouped by CC Phase
```{r}
DimPlot(wat_cc, group.by='Phase', pt.size= 2)
```

Fig. EV1E - WAT cell cycle distribution post-regression
```{r}
cell_counts <- table(wat_cc$wat_unexcluded_clusters, wat_cc$Phase)
print(cell_counts)
```

Fig. EV2A
```{r}
wat_colors=c('#00cd66','#fcd589','#cd6600', '#ff3030','#8c510a','#5ab4ac','#01665e')
DoHeatmap(wat_cc, significant_markers_wat, group.colors= wat_colors, group.by='wat_named_wclusters')+ scale_fill_gradientn(colors = rev(RColorBrewer::brewer.pal(5, "RdBu")))
```

Fig. 1B Pre-exclusion UMAP of WAT
```{r}
#define color scheme
wat_raw_colors= c('#00cd66','#fcd589','#cd6600', '#ff3030','#8c510a','#5ab4ac','#7d26cd','#e066ff')
DimPlot(wat_unexcluded, cols = wat_raw_colors, pt.size=2, group.by='wat_unexcluded_clusters')
```

Fig. 1D Plot violin plots of adipocells and non-adipocell markers 
```{r}
#plot the violins for markers of non-adipocells
Stacked_VlnPlot(wat_unexcluded, colors_use = wat_raw_colors, features= c("Plp1",'Ptprc','Fabp4','Pparg','Dlk1','Pdgfra','Pdgfrb'), x_lab_rotate=45, group.by='wat_unexcluded_clusters')

```


Fig. 1F - Final UMAP of WAT clusters after non-adipocell exclusion and CC regression 
```{r}
wat_colors=c('#00cd66','#fcd589','#cd6600', '#ff3030','#8c510a','#5ab4ac','#01665e')
DimPlot(wat_cc, cols= wat_colors, pt.size=2, group.by='wat_named_wclusters')
```

Fig. 1H - Stacked violin plots of prominent markers
```{r}
Stacked_VlnPlot(wat_cc, features=c("Pdgfra", "Pdgfrb", "Dlk1", "Pparg", "Fabp4", 'Fosl1','Ptchd4','Cpa6','Irx1','Cdca3', 'Kif20a', 'Tgfb1', 'Cxcl14','Serpinb1a','Ly6c1','Cryab','Igfbp2','Gm13889','Bmp4', 'Cd9', 'Npr3'), colors_use=wat_colors, x_lab_rotate = 45, group.by='wat_named_wclusters')
```

Fig. 2D Feature plots 
```{r}
FeaturePlot(wat_cc, "Crlf1", min.cutoff=0.5, max.cutoff = 1.5)
FeaturePlot(wat_cc, "Cryab",min.cutoff=0.8, max.cutoff = 1.5)
FeaturePlot(wat_cc, "Ptn",min.cutoff=2.0, max.cutoff = 4.5)
FeaturePlot(wat_cc, "Ank3",min.cutoff=0.3, max.cutoff = 1.8)
```


Fig. 2F Regulon score patterns
Top panel- CAPs enriched regulons
```{r}
pheatmap(regulon_activity_caps, 
         cluster_cols = TRUE,show_colnames = FALSE,
         cluster_rows = TRUE,
         scale = 'none',  # No scaling, to keep data as is
         breaks=breaks,
         annotation_col = cluster_annotation,
         annotation_colors = list(Cluster = cluster_colors))
```
bottom panel - FAPs enriched regulons
```{r}
pheatmap(regulon_activity_fapls, 
         cluster_cols = TRUE,show_colnames = FALSE,
         cluster_rows = TRUE,
         scale = 'none',  # No scaling, to keep data as is
         breaks=breaks,
         annotation_col = cluster_annotation,
         annotation_colors = list(Cluster = cluster_colors))
```

Table EV1 - wAPC markers
```{r}
write.csv(w1_markers, "w1_markers.csv")
write.csv(w2_markers, "w2_markers.csv")
write.csv(w3_markers, "w3_markers.csv")
write.csv(w4_markers, "w4_markers.csv")
write.csv(wcaps_markers,"wcaps_markers.csv")
write.csv(wfapls_markers, "wfapls_markers.csv")
write.csv(dwfapls_markers, "dwfapls_markers.csv")
```

Table EV3 - DEGs between wCAPs vs wFAPLs
```{r}
wcaps_wfapls_degs <- FindMarkers(wat_cc, ident.1='wCAPs', ident.2 = 'wFAPLs')
write.csv(wcaps_wfapls_degs, file ="DEGs_wCAPs_vs_wFAPLs.csv")
