# Integrated WAT and BAT analysis
This script will generate the following figures:

Figure EV5A - UMAP of aggregated WAT and BAT before regression
Figure EV5B - UMAP of aggregated WAT and BAT before regression, grouped by depot
Figure EV5C -  UMAP of aggregated WAT and BAT before regression, grouped by CC phase
Figure EV5D - Cell count of each cluster in phase before CC regression
Figure EV5E - UMAP of aggregated WAT and BAT after regression, grouped by CC phase
Figure EV5F - Cell count of each cluster in phase before CC regression
Figure EV5G - WAT and BAT cluster distribution in aggregated clusters by cell number count
Figure 4A left -  UMAP of aggreagated WAT and BAT after CC regression
Figure 4A top right -  UMAP of aggreagated WAT and BAT after CC regression, highlighting cells from WAT
Figure 4A bottom right -  UMAP of aggreagated WAT and BAT after CC regression, highlighting cells from BAT
Figure 4B left - Depot composition of WAT BAT aggregate
Figure 4B right and 4C - WAT and BAT cluster distribution in aggregated clusters by percentage
Figure 4D- representative markers of cluster 8 (combined FAPLs)
Figure 4E - Feature plots of depot markers
Figure 4F - co-upregulated genes of wFAPLs and bFAPLs; wCAPs and bCAPs

################
#BEGIN ANALYSIS#
################
## 1. Integrate WAT and BAT
```{r}
wat_bat_nocc <- Reduce(merge, list(wat_cc,bat_cc))
wat_bat_nocc <- SCTransform(wat_bat_nocc)
wat_bat_nocc <- RunPCA(wat_bat_nocc)
wat_bat_nocc <- FindNeighbors(wat_bat_nocc, dims=1:10) 
wat_bat_nocc <- FindClusters(wat_bat_nocc, dims=1:10) #9 clusters
wat_bat_nocc <- RunUMAP(wat_bat_nocc, dims=1:10)

#relabel orig.ident to "White Preadipocytes" and "Brown Preadipocytes
Idents(wat_bat_nocc) <- "orig.ident"
new_cluster_ids <- c('White Preadipocytes', "Brown Preadipocytes")
names(new_cluster_ids) <- levels(wat_bat_nocc)

wat_bat_nocc <- RenameIdents(wat_bat_nocc, new_cluster_ids)
wat_bat_nocc@meta.data$depot<- as.factor(wat_bat_nocc@active.ident)



#relabel numeric seurat clusters to APCs
Idents(wat_bat_nocc) <- "seurat_clusters"
#new_cluster_ids <- c('APCs 1','APCs 2', 'APCs 3', 'APCs 4', 'APCs 5', 'APCs 6', 'APCs 7', 'APCs 8', 'APCs 9') #original clustering 
new_cluster_ids <- c('APCs 1','APCs 2', 'APCs 3', 'APCs 7', 'APCs 6', 'APCs 5', 'APCs 9', 'APCs 4', 'APCs 8') #reordered clustering

names(new_cluster_ids) <- levels(wat_bat_nocc_rds)

wat_bat_nocc <- RenameIdents(wat_bat_nocc_rds, new_cluster_ids)
wat_bat_nocc@meta.data$wat_bat_unregressed_clusters<- as.factor(wat_bat_nocc@active.ident) # new names stored in "wat_bat_unregressed_clusters" metadata
wat_bat_nocc_rds <- PrepSCTFindMarkers(wat_bat_nocc_rds)


DimPlot(wat_bat, pt.size=1.5) #good, this is the plot we want
```


## 2. regressing out cell cycle
```{r}

#cell cycle scoring
library(stringr)
m.s.genes <- str_to_title(cc.genes$s.genes)
m.g2m.genes <- str_to_title(cc.genes$g2m.genes)
wat_bat <- CellCycleScoring(wat_bat, s.features = m.s.genes, g2m.features = m.g2m.genes, set.ident = TRUE)
DimPlot(wat_bat, pt.size=1.5)

#regression starts
wat_bat_cc <- ScaleData(wat_bat, vars.to.regress = c("S.Score", "G2M.Score"), features = rownames(wat_bat))

wat_bat_cc <- RunPCA(wat_bat_cc)
wat_bat_cc <- FindNeighbors(wat_bat_cc, dims=1:10)
wat_bat_cc <- FindClusters(wat_bat_cc, dims=1:10)
wat_bat_cc <- RunUMAP(wat_bat_cc, dims=1:10)
DimPlot(wat_bat_cc, pt.size=1.5)
#DimPlot(wat_bat_cc, pt.size=1.5, group.by= "wat_clusters")
#DimPlot(wat_bat_cc, pt.size=1.5, group.by= "bat_clusters")
```

## 3. Merge the metadata column together for plotting and DEGs
```{r}
# Replace NA values in the original metadata columns with an empty string
wat_bat_cc_rds$wat_named_wclusters[is.na(wat_bat_cc_rds$wat_named_wclusters)] <- ""
wat_bat_cc_rds$bat_named_bclusters[is.na(wat_bat_cc_rds$bat_named_bclusters)] <- ""

# Now, paste the columns together
wat_bat_cc_rds$merged_wat_bat_clusters <- paste(wat_bat_cc_rds$wat_named_wclusters, wat_bat_cc_rds$bat_named_bclusters, sep = "")
DimPlot(wat_bat_cc_rds, group.by="merged_wat_bat_clusters")

```

## 4. rename clusters after regression
```{r}
Idents(wat_bat_cc_rds) <-"seurat_clusters"

new_cluster_ids <- c(1,4,8,5,2,6,7,3,9) #reordered clustering

names(new_cluster_ids) <- levels(wat_bat_cc_rds)

wat_bat_cc_rds <- RenameIdents(wat_bat_cc_rds, new_cluster_ids)
wat_bat_cc_rds@meta.data$wat_bat_regressed_clusters<- as.factor(wat_bat_cc_rds@active.ident) # new names stored in "wat_bat_unregressed_clusters" metadata
```

## 5. Find markers of cluster 8 
```{r}
Idents(wat_bat_cc) <- 'wat_bat_regressed_clusters'
wat_bat_cc <- PrepSCTFindMarkers(wat_bat_cc)
c8_markers <- FindMarkers(wat_bat_cc, ident.1=8)
```

##############################


# FIGURES

## Figure EV5A - UMAP of aggregated WAT and BAT before regression
```{r}
wat_bat_nocc_cols=c("coral","hotpink",'red','brown','tan','navy','plum4','purple','khaki')
DimPlot(wat_bat_nocc, pt.size=2, cols=wat_bat_nocc_cols, group.by= "wat_bat_unregressed_clusters")
```

## Figure EV5B - UMAP of aggregated WAT and BAT before regression, grouped by depot
```{r}
depot_col=c("khaki",'royalblue')
DimPlot(wat_bat_nocc, pt.size=2, cols=depot_col, group.by='depot')
```

## Figure EV5C -  UMAP of aggregated WAT and BAT before regression, grouped by CC phase
```{r}
DimPlot(wat_bat_nocc, pt.size=2, group.by='Phase')
```

## Figure EV5D - Cell count of each cluster in phase before CC regression
```{r}
table(wat_bat_nocc$wat_bat_unregressed_clusters, wat_bat_nocc$Phase)
```

## Figure EV5E - UMAP of aggregated WAT and BAT after regression, grouped by CC phase
```{r}
DimPlot(wat_bat_cc, pt.size=2, group.by='Phase')
```

## Figure EV5F - Cell count of each cluster in phase before CC regression
```{r}
table(wat_bat_cc$wat_bat_regressed_clusters, wat_bat_cc$Phase)
```

## Figure EV5G - WAT and BAT cluster distribution in aggregated clusters by cell number count
```{r}
table(wat_bat_cc$wat_named_wclusters, wat_bat_cc$wat_bat_regressed_clusters)
table(wat_bat_cc$bat_named_bclusters, wat_bat_cc$wat_bat_regressed_clusters)
```


## Figure 4A left -  UMAP of aggreagated WAT and BAT after CC regression
```{r}
combined_colors= c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#C49A00','#a65628','#f781bf','#00868b')
DimPlot(wat_bat_cc, cols=combined_colors, pt.size=2, group.by= 'wat_bat_regressed_clusters')

```

## Figure 4A top right -  UMAP of aggreagated WAT and BAT after CC regression, highlighting cells from WAT
```{r}
depot_col=c("khaki",'white')
DimPlot(wat_bat_cc, pt.size=2, cols=depot_col,group.by='depot')
```


## Figure 4A bottom right -  UMAP of aggreagated WAT and BAT after CC regression, highlighting cells from BAT
```{r}
depot_col=c("white",'royalblue')
DimPlot(wat_bat_cc, pt.size=2, cols=depot_col,group.by='depot')
```


## Figure 4B left - Depot composition of WAT BAT aggregate
```{r}
table(wat_bat_cc$depot, wat_bat_cc$wat_bat_regressed_clusters)
```

## Figure 4B right and 4C - WAT and BAT cluster distribution in aggregated clusters by percentage
```{r}
table(wat_bat_cc$wat_named_wclusters, wat_bat_cc$wat_bat_regressed_clusters)
table(wat_bat_cc$bat_named_bclusters, wat_bat_cc$wat_bat_regressed_clusters)

#tables were saved then calculated percentage in excel
```

## Figure 4D- representative markers of cluster 8 (combined FAPLs)
```{r}
combined_colors= c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#C49A00','#a65628','#f781bf','#00868b')
#names(combined_colors)= c(1,2,3,4,5,6,7,8,9)

Stacked_VlnPlot(wat_bat_cc, features= c('Col4a2','Col4a1','Cryab','Meox2','Serpinb1a',"Nr2f2",'Pparg'), colors_use=combined_colors, )
```

## Figure 4E - Feature plots of depot markers
```{r}
FeaturePlot(wat_bat_cc, c("Hoxc10","Hoxc9","Zic1")) #feature plots

Idents(wat_bat_cc) = "merged_wat_bat_clusters"
Stacked_VlnPlot(wat_bat_cc, features= c("Hoxc10","Hoxc9","Zic1"), idents = c('wFAPLs', 'bFAPLs'), colors_use = c("#00868b","#FF0000") ) #Violin plot of bFAPLs vs wFAPLs
```


#Figure 4F - co-upregulated genes of wFAPLs and bFAPLs; wCAPs and bCAPs
```{r}
#need wat_cc and bat_cc object loaded
DefaultAssay(bat_cc) <- "RNA"
DefaultAssay(wat_cc) <- "RNA"
bfapls_markers= FindMarkers(bat_cc, ident.1='bFAPLs') 
wfapls_markers= FindMarkers(wat_cc, ident.1='wFAPLs')
bcaps_markers= FindMarkers(bat_cc, ident.1='bCAPs')
wcaps_markers= FindMarkers(wat_cc, ident.1='wCAPs')

bfapls_markers.up= bfapls_markers %>%
  filter(p_val_adj < 0.05 & avg_log2FC>0)
wfapls_markers.up= wfapls_markers %>%
  filter(p_val_adj < 0.05 & avg_log2FC>0)
bcaps_markers.up= bcaps_markers %>%
  filter(p_val_adj < 0.05 & avg_log2FC>0)
wcaps_markers.up= wcaps_markers %>%
  filter(p_val_adj < 0.05 & avg_log2FC>0)

common_caps= intersect(rownames(bcaps_markers.up), rownames(wcaps_markers.up))
common_fapls= intersect(rownames(bfapls_markers.up), rownames(wfapls_markers.up))
#then count the number in each list. 

#note: rerunning findmarkers can produce different results if some packages are updated.

#originally, the data was saved and processed on excel
#then, the gene names were separated on comparetwolists.com, where:
#list A: wcaps or wfapls up markers (p_val_adj<0.05)
#list B: bcaps or bfapls up markers (p_val_adj<0.05)
#the output will show genes that are shared and unique

#################################
wat <- SCTransform(wat_cc)
DefaultAssay(wat_cc)
wcaps_markers= FindMarkers(wat_cc,'wCAPs') %>% filter(p_val_adj<0.05)
wfapls_markers= FindMarkers(wat_cc, 'wFAPLs') %>% filter(p_val_adj<0.05)

DefaultAssay(bat_cc)
Idents(bat_cc) <- 'bat_named_bclusters'
bcaps_markers= FindMarkers(bat,'bCAPs')%>% filter(p_val_adj<0.05) #for BAT
bfapls_markers= FindMarkers(bat, 'bFAPLs')%>% filter(p_val_adj<0.05)

#for the unfiltered pct.1 venn diagram

sc_wfapls <- wfapls_markers %>% filter(p_val_adj<0.05, avg_log2FC>0) %>% rownames()
sc_bfapls <- bfapls_markers %>% filter(p_val_adj<0.05, avg_log2FC>0) %>% rownames()
venn.plot <- venn.diagram(
  x = list(Sc_wFAPLs = sc_wfapls, Sc_bFAPLs = sc_bfapls),
  filename = NULL,
  fill = c("khaki", "royalblue"), alpha = 0.5
)
grid::grid.draw(venn.plot)



sc_wcaps <- wcaps_markers %>% filter(p_val_adj<0.05, avg_log2FC>0) %>% rownames()
sc_bcaps <- bcaps_markers %>% filter(p_val_adj<0.05, avg_log2FC>0) %>% rownames()
venn.plot <- venn.diagram(
  x = list(Sc_wCAPs = sc_wcaps, Sc_bCAPs = sc_bcaps),
  filename = NULL,
  fill = c("khaki", "royalblue"), alpha = 0.5
)
grid::grid.draw(venn.plot)


#get list of unique and overlap unfiltered pct
shared_sc_fapl_markers <- intersect(sc_wfapls, sc_bfapls)
unique_sc_wfapls <- setdiff(sc_wfapls, sc_bfapls)
unique_sc_bfapls <- setdiff(sc_bfapls, sc_wfapls)

shared_sc_caps_markers <- intersect(sc_wcaps, sc_bcaps)
unique_sc_wcaps <- setdiff(sc_wcaps, sc_bcaps)
unique_sc_bcaps <- setdiff(sc_bcaps, sc_wcaps)

#filter for .75 pct genes
sc_wcaps_0.75 <- wcaps_markers %>% filter(p_val_adj<0.05, avg_log2FC>0, pct.1>0.75) %>% rownames()
sc_bcaps_0.75 <- bcaps_markers %>% filter(p_val_adj<0.05, avg_log2FC>0, pct.1>0.75) %>% rownames()

sc_wfapls_0.75 <- wfapls_markers %>% filter(p_val_adj<0.05, avg_log2FC>0, pct.1>0.75) %>% rownames()
sc_bfapls_0.75 <- bfapls_markers %>% filter(p_val_adj<0.05, avg_log2FC>0, pct.1>0.75) %>% rownames()

#get list of unique and overlap filtered pct.1>0.75
shared_0.75_sc_fapl_markers <- intersect(sc_wfapls, sc_bfapls)
unique_0.75_sc_wfapls <- setdiff(sc_wfapls, sc_bfapls)
unique_sc_0.75_bfapls <- setdiff(sc_bfapls, sc_wfapls)

shared_0.75_sc_caps_markers <- intersect(sc_wcaps_0.75, sc_bcaps_0.75)
unique_0.75_sc_wcaps <- setdiff(sc_wcaps_0.75, sc_bcaps_0.75)
unique_0.75_sc_bcaps <- setdiff(sc_bcaps_0.75, sc_wcaps_0.75)


```

#####################
###END OF ANALYSIS###
#####################



save all objects for reproducibility
```{r}
bat_nocc$bat_unexcluded_clusters <- bat_unexcluded$bat_unexcluded_clusters
bat_cc$bat_unexcluded_clusters <- bat_unexcluded$bat_unexcluded_clusters
```

```{r}
wat_nocc$wat_unexcluded_clusters <- wat_unexcluded$wat_unexcluded_clusters
wat_cc$wat_unexcluded_clusters <- wat_unexcluded$wat_unexcluded_clusters
```

```{r}
wat_bat_cc_diet <- DietSeurat(object = wat_bat_cc_rds,
  layers =c("data", "scale.data") ,
  assays=c("RNA", "AUCell", "SCT"),
  data = TRUE,     # Keep the normalized data slot (for FeaturePlot)
  counts = NULL,  # Discard raw counts to save space
  scale.data = TRUE, # Discard scaled data to save space (unless you need it for specific reasons)
  dimreducs = c("pca", "umap"), # Keep PCA and UMAP embeddings
  graphs = FALSE, # Discard graphs (unless specific graph analysis is central to your paper)
  misc = FALSE, # Discard misc slot (often not critical)
)
wat_bat_cc_diet[["RNA"]]["scale.data"] <- NULL
wat_bat_cc_diet[["RNA"]]["counts"] <- NULL
wat_bat_cc_diet[["SCT"]]["counts"] <- NULL
wat_bat_cc_diet[["SCT"]]["scale.data"] <- NULL

wat_bat_cc_diet@assays$RNA@layers$counts.0watpreads <- NULL
wat_bat_cc_diet@assays$RNA@layers$counts.0BATpreads <- NULL

wat_bat_cc_diet@assays$RNA <- NULL
wat_bat_cc_diet@assays$RNA <- NULL

#wat_bat_nocc_diet@assays$RNA@layers$counts.0watpreads <- NULL
#wat_bat_nocc_diet@assays$RNA@layers$counts.0BATpreads <- NULL



saveRDS(wat_bat_cc_diet, file="F:/new_analysis/scripts/publication/wat_bat_cc.rds")
saveRDS(wat_bat_cc_rds, file="F:/new_analysis/scripts/publication/wat_bat_cc_FULL.rds")
```

