# bat PROCESSING
This script will generate the following figures:
Fig. EV2A- bat UMAP without non-adipocells, grouped by original clusterings
Fig. EV2B- bat UMAP without non-adipocells before CC regression, grouped by CC Phase
Fig. EV2C- bat cell cycle distribution pre-regression
Fig. EV2D - bat UMAP without non-adipocells after CC regression, grouped by CC Phase
Fig. EV2E - bat cell cycle distribution post-regression
Fig. EV2B - markers of top 10% wAPC heatmap
Fig. 1C Pre-exclusion UMAP of bat
Fig. 1E Plot violin plots of adipocells and non-adipocell markers 
Fig. 1G - Final UMAP of bat clusters after non-adipocell exclusion and CC regression 
Fig. 1I - Stacked violin plots of prominent markers
Fig. 3D Feature plots 
Fig. 3F Regulon score patterns
Table EV2 - wAPC markers
Table EV4 - DEGs between wCAPs vs wFAPLs

#A. Seurat processing

## 1. Load in libraries
```{r}
library(Seurat)
library(readr)
library(dplyr)
library(patchwork)
library(data.table)
library(ggplot2)
library(cowplot)
library(pheatmap)
library(sctransform)
library(SeuratDisk)
library(scater)
library(tidyverse)
library(SeuratWrappers)
library(gplots)
library(stringr)
library(scCustomize)
```

## 2. Load in data
```{r}
bat.data = Read10X(data.dir = "//research.drive.wisc.edu/galmozzi/Galmozzi lab/Hoang/Projects/Adipocyte Subtypes/new_analysis/0BAT/filtered_feature_bc_matrix")

```

```{r}
#convert into seurat object
bat = CreateSeuratObject(counts = bat.data, project = "0BATpreads", min.cells = 3, min.features = 200)
bat$data.set = rep("bat", length(bat$orig.ident))
```

## 3. Calculate % mitochondria RNA and apply filters
```{r}
#calculate percent mitochondrial RNA for batpreads
bat[["percent.mt"]] = PercentageFeatureSet(bat, pattern = "^mt-")

#generate violin plot #genes/cell, mtRNA/cell, reads/cell
VlnPlot(bat, features= c("nFeature_RNA", "nCount_RNA", "percent.mt"), pt.size=0.5, ncol = 3)

#apply filter-- UMI 6000-60000, features 3900-7700, mito umi 1.7-5
bat <- subset(bat, subset = nFeature_RNA > 3900 & nFeature_RNA < 7700 & percent.mt < 5& percent.mt > 1.7 & nCount_RNA > 6000 & nCount_RNA < 60000) 
```

## 4. Data normalization and scaling
```{r}
#normalizing the data -normalizes the feature expression measurements for each cell by the total expression, multiplies this by a scale factor (10,000 by default), and log-transforms the result
bat <- NormalizeData(bat, normalization.method = "LogNormalize", scale.factor = 10000)

#identification of highly variable features
bat <- FindVariableFeatures(bat, selection.method = "vst", nfeatures = 2000)

#scaling the data 
all.genes <- rownames(bat)
bat <- ScaleData(bat, features = all.genes)
```

## 5. Run PCA and UMAP clustering
```{r}
#PCA
bat <- RunPCA(bat, features = VariableFeatures(object = bat))
#clustering
bat <- FindNeighbors(bat, dims = 1:10)
bat <- FindClusters(bat, resolution = 0.5)
#umap
bat <- RunUMAP(bat, dims = 1:10)
DimPlot(bat, reduction = "umap")
```

## 6. Find all markers to identify non-adipocells

```{r}
#find all markers
bat.markers <- FindAllMarkers(bat,features = VariableFeatures(object = bat))

#renaming identities for subclusters based on the markers identified
new_cluster_ids <- c('APCs 1', 'APCs 2', "APCs 3", "APCs 4",'APCs 5', 'APCs 6', 'APCs 7',  "Myeloid cells", 'Glial cells' , 'Myoblasts')
names(new_cluster_ids) <- levels(bat)
bat <- RenameIdents(bat, new_cluster_ids)

#create a new metadata column called "bat_unexcluded_clusters"
bat@meta.data$bat_unexcluded_clusters <- as.factor(bat@active.ident) 
```

Save clustering information
```{r}
write.csv(bat$bat_unexcluded_clusters, "F:/new_analysis/scripts/publication/bat_unexcluded_clusters.csv")
write.csv(bat@reductions$pca@cell.embeddings, "F:/new_analysis/scripts/publication/bat_unexcluded_pca.csv")
write.csv(bat@reductions$pca@feature.loadings, "F:/new_analysis/scripts/publication/bat_unexcluded_pca_loadings.csv")
write.csv(bat@reductions$umap@cell.embeddings, "F:/new_analysis/scripts/publication/bat_unexcluded_umap.csv")
```


## 7. Remove non-adipocells and recluster
```{r}
#redefine the bat object as bat_unexclude for future reference
bat_unexcluded <- bat

bat_excluded <- subset(x=bat_unexcluded, 
                      idents = c("Glial cells", "Myeloid cells", "Myoblasts"),
                      invert = TRUE) #remove glial and myeloid cells

bat_excluded <- NormalizeData(bat_excluded, normalization.method = "LogNormalize", scale.factor = 10000)
bat_excluded <- FindVariableFeatures(bat_excluded, selection.method = "vst")
bat_excluded <- RunPCA(bat_excluded, features = VariableFeatures(object = bat_excluded))
bat_excluded <- FindNeighbors(bat_excluded, dims = 1:10)
bat_excluded <- FindClusters(bat_excluded, resolution = 0.5)
bat_excluded <- RunUMAP(bat_excluded, dims = 1:10)
```


## 8. Cell cycle regression
## Cell cycle scoring
```{r}
#rename bat_excluded to bat_nocc to track cell cycle status
bat_nocc = bat_excluded

#cell cycle scoring
m.s.genes <- str_to_title(cc.genes$s.genes)
m.g2m.genes <- str_to_title(cc.genes$g2m.genes)
bat_nocc <- CellCycleScoring(bat_nocc, s.features = m.s.genes, g2m.features = m.g2m.genes, set.ident = FALSE)
```
## Cell cycle regression
```{r}
#REGRESS OUT CELL CYCLE SCORES DURING DATA SCALING
bat_cc <- ScaleData(bat_nocc, vars.to.regress = c("S.Score", "G2M.Score"), features = rownames(bat_nocc))
#run PCA 
bat_cc <- RunPCA(bat_cc, features=VariableFeatures(bat_cc), nfeatures.print=10)
#RUN UMAP ON CC 
bat_cc <- FindNeighbors(bat_cc, dims = 1:10)
bat_cc<- FindClusters(bat_cc, resolution = 0.5)
bat_cc <- RunUMAP(bat_cc, dims = 1:10, return.model=TRUE)
```


## 9. final cluster names
```{r}
Idents(bat_cc) <- "seurat_clusters"
new_cluster_ids <- c('Early Preads B1', 'Early Preads B2', "bFAPLs", "bCAPs",'Early Preads B3', 'Early Preads B4', 'Early Preads B5')
names(new_cluster_ids) <- levels(bat_cc)
bat_cc <- RenameIdents(bat_cc, new_cluster_ids)
bat_cc@meta.data$bat_named_wclusters <- as.factor(bat_cc@active.ident)



bat_cc$bat_named_bclusters <- factor(bat_cc$bat_named_bclusters, levels = c('Early Preads B1', "Early Preads B2", "Early Preads B3", "Early Preads B4","Early Preads B5","bCAPs",'bFAPLs')
Idents(bat_cc) <- "bat_named_bclusters"
```



## 10. Find markers of each cluster
```{r}
#bat_cc <- SCTransform(bat_cc)
bcaps_markers= FindMarkers(bat_cc_diet,'bCAPs') #for BAT
bfapls_markers= FindMarkers(bat_cc_diet, 'bFAPLs')
b1_markers = FindMarkers(bat_cc_diet, 'Early Preads B1')
b2_markers= FindMarkers(bat_cc_diet, 'Early Preads B2')
b3_markers= FindMarkers(bat_cc_diet, 'Early Preads B3')
b4_markers= FindMarkers(bat_cc_diet, 'Early Preads B4')
b5_markers= FindMarkers(bat_cc_diet, 'Early Preads B5')
```

Filter for top 10% markers
```{r}
b1_markers_filtered <- b1_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.75)  %>%
  arrange(desc(avg_log2FC)) %>%  # Sort wy avg_log2FC in descending order
    slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC   # Select the top 50 genes
b2_markers_filtered <- b2_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.75) %>%
  arrange(desc(avg_log2FC)) %>%  
    slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC   
b3_markers_filtered <- b3_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.75) %>%
  arrange(desc(avg_log2FC)) %>%  
    slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC   
b4_markers_filtered <- b4_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.75) %>%
  arrange(desc(avg_log2FC)) %>%  
    slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC 
b5_markers_filtered <- b5_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.75) %>%
  arrange(desc(avg_log2FC)) %>%  
    slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC   
bcaps_markers_filtered <- bcaps_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.5) %>%
  arrange(desc(avg_log2FC)) %>%  
    slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC   
bfapls_markers_filtered <- bfapls_markers %>%
  filter(p_val_adj < 0.05 & pct.1>0.75) %>%
  arrange(desc(avg_log2FC)) %>%  
    slice_max(order_by = avg_log2FC, prop = 0.10) #alternatively take top 10% by L2FC   


#extract just the name of genes
bfapls_markers_list <-rownames(bfapls_markers_filtered)
bcaps_markers_list <- rownames(bcaps_markers_filtered)
bfapls_markers_list <- rownames(bfapls_markers_filtered)
b1_markers_list <-rownames(b1_markers_filtered)
b2_markers_list <- rownames(b2_markers_filtered)
b3_markers_list <- rownames(b3_markers_filtered)
b4_markers_list <- rownames(b4_markers_filtered)
b5_markers_list <- rownames(b5_markers_filtered)

#concatenate all significant markers into 1 single vector
significant_markers_bat= c(b1_markers_list,b2_markers_list,b3_markers_list,b4_markers_list, b5_markers_list, bcaps_markers_list,bfapls_markers_list)
```


## 11. Export UMAP and metadata for RNA velocity analysis
```{r}
#export uMAP projections
umap = cbind("Barcode" = rownames(Embeddings(object = bat_cc, reduction = "umap")), Embeddings(object = bat_cc, reduction = "umap"))
write.table(umap, file="F:/uMAP_seurat.csv", sep = ",", quote = F, row.names = F, col.names = T)

#export metadata to get clusters export
write.csv(bat_cc@meta.data,"F:/seurat_metadata.csv") #use excel to extract barcodes and cluster names later

```


# B. Scenic analysis
Here, we will be using the aucell score scenic output for bat, merging the matrix with the bat_cc object and perform DE expression for AUCell differential scores.

## 1. Load in libraries
```{r}
library(Seurat)
library(dplyr)
```

## 2. Load in data
```{r}
#bat <- readRDS('F:/new_analysis/labeled_analysis/bat.rds')
aucell_scores <- read.csv("F:/scenic/bat/data/auc_mtx_v10.csv", row.names = 1)

seurat_object=bat_cc_diet #set current object to be bat
```

## 3. Process the AUCell matrix
```{r}
# Cells should be columns, regulons should be rows
if (!identical(colnames(aucell_scores), colnames(seurat_object))) {
  aucell_scores <- t(aucell_scores)
}
# Set proper names if needed
# Create a new assay object with AUCell scores
aucell_assay <- CreateAssayObject(counts = aucell_scores)

# Add this new assay to your Seurat object
seurat_object[["AUCell"]] <- aucell_assay
DefaultAssay(seurat_object) <- "AUCell"
seurat_object<- ScaleData(seurat_object)
```

## 4. Find differential scores
```{r}
Idents(seurat_object) <- "bat_named_bclusters"
bcaps_vs_bfapls_tcf <- FindMarkers(seurat_object, ident.1='bCAPs', ident.2='bFAPLs')
bcaps_vs_bfapls_tcf_filtered <- bcaps_vs_bfapls_tcf %>%
  filter(p_val_adj < 0.05 & avg_log2FC>0& pct.1>0.5)

bcaps_tcf_list <- rownames(bcaps_vs_bfapls_tcf_filtered)

bfapls_vs_bcaps_tcf <- FindMarkers(seurat_object, ident.1='bFAPLs', ident.2='bCAPs')
bfapls_vs_bcaps_tcf_filtered <- bfapls_vs_bcaps_tcf %>%
  filter(p_val_adj < 0.05 & avg_log2FC>0 & pct.1>0.5)
bfapls_tcf_list <- rownames(bfapls_vs_bcaps_tcf_filtered)

```

## 5. Filter out TCF that shows erratic expressions
```{r}
#FINAL HEATMAPS 3.5.25
bcaps_tcf_to_remove <- c('Stat5b...') #adipogenic
bfapls_tcf_to_remove <- c('E2f2...','Hoxb2...','Irf8...','Gata3...','Rfxank...', 'Hoxb7...', 'Shox2...') #final list 3/5/25

# Remove the genes from the list
adipogenic_tcf= bcaps_tcf_list
adipogenic_tcf_cleaned <- adipogenic_tcf[!adipogenic_tcf %in% bcaps_tcf_to_remove] 


fibrogenic_tcf= bfapls_tcf_list
fibrogenic_tcf_cleaned <- fibrogenic_tcf[!fibrogenic_tcf %in% bfapls_tcf_to_remove] 

```

## 6. subset data for heatmap plotting
```{r}
#extract the data matrix from seurat object
data_matrix <- GetAssayData(object = seurat_object, assay = "AUCell", slot='scale.data') 
#setting up the labels and colors
cluster_labels <- seurat_object$bat_named_bclusters
# Order data matrix by cluster labels
ordered_indices <- order(cluster_labels)
ordered_data_matrix <- data_matrix[, ordered_indices]
ordered_cluster_labels <- cluster_labels[ordered_indices]
# Create annotation data frame
cluster_annotation <- data.frame(Cluster = factor(ordered_cluster_labels))

#define color for each cluster
cluster_colors<- setNames(c("#00868b", "#cd5b45", "#FFE300","#00BFFF","#00FFFF","#0000FF", "#FF0000"), sort(unique(cluster_labels)))
# Create the annotation data frame
cluster_annotation <- data.frame(Cluster = factor(ordered_cluster_labels, levels = names(cluster_colors)))


#Set limits and breaks based on standard deviation
# Calculate the standard deviation of the matrix
std_dev <- sd(as.vector(data_matrix))
# Define breaks around zero, within ±2 standard deviations
max_limit <- 2 * std_dev
breaks <- seq(-max_limit, max_limit, length.out = 101)
```

Select CAPs or FAPs heatmap
```{r}
regulon_activity_caps <- ordered_data_matrix[adipogenic_tcf_cleaned, ] #for CAPs heatmap
regulon_activity_fapls <- ordered_data_matrix[fibrogenic_tcf_cleaned, ] # for FAPLs heatmap

```

Plot the heatmap
```{r}
pheatmap(regulon_activity_caps, #or regulon_activity_fapls
         cluster_cols = TRUE,show_colnames = FALSE,
         cluster_rows = TRUE,
         scale = 'none',  # No scaling, to keep data as is
         breaks=breaks,
         annotation_col = cluster_annotation,
         annotation_colors = list(Cluster = cluster_colors))
```





# FIGURE PLOTS

Fig. EV1F- bat UMAP without non-adipocells, grouped by original clusterings
```{r}
#define color scheme
bat_raw_colors= c("#00868b", "#cd5b45", "#FFE300","#00BFFF","#00FFFF","#0000FF", "#FF0000", '#e066ff','#7d26cd','#FF69B4')
DimPlot(bat_nocc, reduction = "umap", cols=bat_raw_colors, group.by = "bat_unexcluded_clusters", pt.size=2)
```

Fig. EV1G- bat UMAP without non-adipocells before CC regression, grouped by CC Phase
```{r}
DimPlot(bat_nocc, reduction = "umap", group.by = "Phase", pt.size=2)
```

Fig. EV1H- bat cell cycle distribution pre-regression
```{r}
cell_counts <- table(bat_nocc$bat_unexcluded_clusters, bat_nocc$Phase)
print(cell_counts)

```

Fig. EV1I - bat UMAP without non-adipocells after CC regression, grouped by CC Phase
```{r}
DimPlot(bat_cc, group.by='Phase', pt.size= 2)
```

Fig. EV1J - bat cell cycle distribution post-regression
```{r}
cell_counts <- table(bat_cc$bat_named_bclusters, bat_cc$Phase)
print(cell_counts)
```

Fig. EV2B - heatmap of top 10% markers per cluster
```{r}
bat_colors= c("#00868b", "#cd5b45", "#FFE300","#00BFFF","#00FFFF","#0000FF", "#FF0000")
DoHeatmap(bat_cc_diet, significant_markers_bat, group.colors= bat_colors)+ scale_fill_gradientn(colors = rev(RColorBrewer::brewer.pal(5, "RdBu")))
```

Fig. 1B Pre-exclusion UMAP of bat
```{r}
DimPlot(bat_unexcluded, cols = bat_raw_colors, pt.size=2, group.by="bat_unexcluded_clusters")
```

Fig. 1D Plot violin plots of adipocells and non-adipocell markers 
```{r}
#plot the violins for markers of non-adipocells
Stacked_VlnPlot(bat_unexcluded, colors_use = bat_raw_colors, features= c("Myod1","Plp1",'Ptprc','Fabp4','Pparg','Dlk1','Pdgfra','Pdgfrb'), x_lab_rotate=45)

```


Fig. 1F - Final UMAP of bat clusters after non-adipocell exclusion and CC regression 
```{r}
bat_colors= c("#00868b", "#cd5b45", "#FFE300","#00BFFF","#00FFFF","#0000FF", "#FF0000") 
DimPlot(bat_cc, cols= bat_colors, pt.size=2)
```

Fig. 1H - Stacked violin plots of prominent markers
```{r}
Stacked_VlnPlot(bat_cc, features=c("Pdgfra", "Pdgfrb", "Dlk1", "Pparg", "Fabp4", 'Sema3a','Nav3','Lrrc15','Cdca3','Kif20a','Creb5','Ptprz1','Igf1','Hp','Mecom','Serpinb1a','Cryab'), colors_use=bat_colors, x_lab_rotate = 45)

```

Fig. 3D Feature plots 
```{r}
DefaultAssay(bat_cc) <- 'RNA'
FeaturePlot(bat_cc, "Crlf1", min.cutoff=0.5, max.cutoff = 2)
FeaturePlot(bat_cc, "Cryab",min.cutoff=0.5, max.cutoff = 2)
FeaturePlot(bat_cc, "Ptn",min.cutoff=1.5, max.cutoff = 3.5)
FeaturePlot(bat_cc, "Ank3",min.cutoff=0.5, max.cutoff = 1.5)
```


Fig. 3F Regulon score patterns
Top panel- CAPs enriched regulons
```{r}
pheatmap(regulon_activity_caps, 
         cluster_cols = TRUE,show_colnames = FALSE,
         cluster_rows = TRUE,
         scale = 'none',  # No scaling, to keep data as is
         breaks=breaks,
         annotation_col = cluster_annotation,
         annotation_colors = list(Cluster = cluster_colors))
```
bottom panel - FAPs enriched regulons
```{r}
pheatmap(regulon_activity_fapls, 
         cluster_cols = TRUE,show_colnames = FALSE,
         cluster_rows = TRUE,
         scale = 'none',  # No scaling, to keep data as is
         breaks=breaks,
         annotation_col = cluster_annotation,
         annotation_colors = list(Cluster = cluster_colors))
```

Table EV2 - bAPC markers
```{r}
write.csv(b1_markers, "b1_markers.csv")
write.csv(b2_markers, "b2_markers.csv")
write.csv(b3_markers, "b3_markers.csv")
write.csv(b4_markers, "b4_markers.csv")
write.csv(b5_markers, "b4_markers.csv")
write.csv(bcaps_markers,"bcaps_markers.csv")
write.csv(bfapls_markers, "bfapls_markers.csv")

```

Table EV4 - DEGs between bCAPs vs bFAPLs
```{r}
wcaps_wfapls_degs <- FindMarkers(bat_cc, ident.1='wCAPs', ident.2 = 'wFAPLs')
write.csv(wcaps_wfapls_degs, file ="DEGs_wCAPs_vs_wFAPLs.csv")
```

#############################
########END OF ANALYSIS######
#############################

save all objects for reproducibility
```{r}
bat_nocc$bat_unexcluded_clusters <- bat_unexcluded$bat_unexcluded_clusters
bat_cc$bat_unexcluded_clusters <- bat_unexcluded$bat_unexcluded_clusters
```

```{r}
wat_nocc$wat_unexcluded_clusters <- wat_unexcluded$wat_unexcluded_clusters
wat_cc$wat_unexcluded_clusters <- wat_unexcluded$wat_unexcluded_clusters
```

```{r}
wat_nocc_diet <- DietSeurat(object = wat_nocc,
  layers =c("data", "scale.data") ,
  assays=c("RNA", "AUCell", "SCT"),
  data = TRUE,     # Keep the normalized data slot (for FeaturePlot)
  counts = NULL,  # Discard raw counts to save space
  scale.data = TRUE, # Discard scaled data to save space (unless you need it for specific reasons)
  dimreducs = c("pca", "umap"), # Keep PCA and UMAP embeddings
  graphs = FALSE, # Discard graphs (unless specific graph analysis is central to your paper)
  misc = FALSE, # Discard misc slot (often not critical)
)
wat_nocc_diet[["RNA"]]["scale.data"] <- NULL
wat_nocc_diet[["RNA"]]["counts"] <- NULL
wat_nocc_diet[["SCT"]]["counts"] <- NULL
wat_nocc_diet[["SCT"]]["scale.data"] <- NULL
saveRDS(wat_nocc_diet, file="F:/new_analysis/scripts/publication/wat_nocc.rds")
```

```{r}
seurat_object= wat_unexcluded

Idents(seurat_object) = "wat_named_clusters"
#new_cluster_ids <- c('APCs 1', 'APCs 2', "APCs 3", "APCs 4",'APCs 5', 'APCs 6', 'APCs 7',  "Myeloid cells", 'Glial cells' , 'Myoblasts')

new_cluster_ids <- c('APCs 1', 'APCs 2', "APCs 3", "APCs 4",'APCs 5', 'APCs 6',  'Glial cells' , "Myeloid cells")

names(new_cluster_ids) <- levels(seurat_object)
seurat_object <- RenameIdents(seurat_object, new_cluster_ids)

#create a new metadata column called "bat_unexcluded_clusters"
#seurat_object@meta.data$bat_unexcluded_clusters <- as.factor(seurat_object@active.ident) 

seurat_object@meta.data$wat_unexcluded_clusters <- as.factor(seurat_object@active.ident) 

wat_unexcluded = seurat_object

```

