# BAT Reference mapping for external datasets

This script will generate the following figures:
Fig. 5C - Reference mapping of 2 datasets (karlina, burl)
Fig. 5D - quantification of labeled cells from mapping


## Data downloaded links:
Karlina - https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE161447
Burl - https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE207707
https://github.com/RBBurl1227/eLife-2022-ColdInducedBrownAdipocyteNeogenesis


# Part A- process external data

### libraries
```{r}
library(Seurat)
library(SeuratData)
library(glmGampoi)
library(Seurat.utils)
library(scDblFinder)
library(dittoSeq)
library(pheatmap)
library(gprofiler2)

library(tximport)
library(AnnotationHub)
library(edgeR)
library(AnnotationDbi)

library(RColorBrewer)
library(tidyverse)
library(ggplot2)
library(dplyr)
```

### karlina dataset
```{r}
#read in data
karlina.data = Read10X(data.dir = "F:/new_analysis/integrated_analysis/bat/Karlina2021")

#convert into seurat object
karlina = CreateSeuratObject(counts = karlina.data, project = "karlina", min.cells = 3, min.features = 200)
karlina = CreateSeuratObject(counts = karlina.data, project = "karlina", min.cells = 3, min.features = 200)
karlina$data.set = rep("karlina", length(karlina$orig.ident))

#calculate percent mitochondrial RNA for karlinapreads
karlina[["percent.mt"]] = PercentageFeatureSet(karlina, pattern = "^mt-")
karlina[["percent.mt"]] = PercentageFeatureSet(karlina, pattern = "^MT-")

karlina[["percent.ribo"]] <- PercentageFeatureSet(karlina, pattern="^rps")


karlina[["percent.ribo"]] <- PercentageFeatureSet(karlina, features = ribo.genes)
####################end##########
#generate violin plot #genes/cell, mtRNA/cell, reads/cell
VlnPlot(karlina, features= c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.ribo"), pt.size=0.5, ncol = 3)
#rerun without the dots
VlnPlot(karlina, features= c("nFeature_RNA", "nCount_RNA", "percent.mt"), pt.size=0, ncol = 3)

#Feature Scatter visualize feature-feature relationships, but can be used
# for anything calculated by the object, i.e. columns in object metadata, PC scores etc.
plot1 <- FeatureScatter(karlina, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(karlina, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2



#apply filter-- UMI 6000-60000, features 3900-7700, mito umi 1.7-5
karlina <- subset(karlina, subset = nCount_RNA<15000 & percent.mt < 20)
karlina<- UpdateSeuratObject(karlina)

#normalizing the data -normalizes the feature expression measurements for each cell by the total expression, multiplies this by a scale factor (10,000 by default), and log-transforms the result
karlina <- NormalizeData(karlina, normalization.method = "LogNormalize", scale.factor = 10000)

#identification of highly variable features
karlina <- FindVariableFeatures(karlina, selection.method = "vst", nfeatures = 4000)

#Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(karlina), 10)
top10
# plot variable features with and without labels
plot1 <- VariableFeaturePlot(wat)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2

#scaling the data -pre-processing step prior to dimensional reduction techniques like PCA. The ScaleData() function:
#Shifts the expression of each gene, so that the mean expression across cells is 0
#Scales the expression of each gene, so that the variance across cells is 1
#This step gives equal weight in downstream analyses, so that highly-expressed genes do not dominate
#The results of this are stored in pbmc[["RNA"]]@scale.data 
all.genes <- rownames(karlina)
karlina <- ScaleData(karlina, features = all.genes)

karlina <-SCTransform(karlina)



#PCA
karlina <- RunPCA(karlina, features = VariableFeatures(object = karlina))

#determine dimensionality
ElbowPlot(karlina, ndims = 50, reduction = "pca")
karlina <- JackStraw(karlina, num.replicate=100)
karlina <- ScoreJackStraw(karlina, dims = 1:20)
JackStrawPlot(karlina, dims = 1:15)

#clustering
karlina <- FindNeighbors(karlina, dims = 1:50)
karlina <- FindClusters(karlina, resolution = 0.5)
karlina <- FindClusters(karlina, resolution = 0.25)

#umap
karlina <- RunUMAP(karlina, dims = 1:50)
DimPlot(karlina)

#renaming clusters
new_cluster_names <- c("0" = "APC1",
                       "3" = "APC2",
                       "5" = "Mature BAT")
# Rename clusters in the metadata
karlina <- RenameIdents(karlina, new_cluster_names)
Idents(karlina.preads) <- seurat_object$new_metadata_column

#########################################################
###############SUBSETTING ONLY ADIPOCELLS################
#########################################################


karlina.preads <-subset(karlina, idents=c("APC1", "APC2"), invert=FALSE)
karlina.preads <-subset(karlina, idents=c(0,8), invert=FALSE)
karlina.preads <- SCTransform(karlina.preads)


karlina.preads <- NormalizeData(karlina.preads, normalization.method = "LogNormalize", scale.factor = 10000)
#identification of highly variable features
karlina.preads <- FindVariableFeatures(karlina.preads, selection.method = "vst", nfeatures = 4000)  
all.genes <- rownames(karlina.preads)
karlina.preads <- ScaleData(karlina.preads, features = all.genes)

  
  
#change names of clusters to "B"+number
karlina.preads@meta.data$karlina_clusters <- as.factor(karlina.preads@active.ident)

#PCA
karlina.preads <- RunPCA(karlina.preads, features = VariableFeatures(object = karlina.preads))

#determine dimensionality
ElbowPlot(karlina.preads, ndims = 50, reduction = "pca")
karlina.preads <- JackStraw(karlina.preads, num.replicate=100)
karlina.preads <- ScoreJackStraw(karlina.preads, dims = 1:20)
JackStrawPlot(karlina.preads, dims = 1:15)

#clustering
karlina.preads <- FindNeighbors(karlina.preads, dims = 1:50)
karlina.preads <- FindClusters(karlina.preads, resolution = 0.7)

#umap
karlina.preads <- RunUMAP(karlina.preads, dims = 1:50)
DimPlot(karlina.preads, pt.size=2)
```

### Burl dataset
```{r}
#according to how they constructed their data, they first merged the RT and COLD in each condition using merge() for each rep
#then they used FindIntegrationAnchors and IntegrateData on 2 replicates
#So, my plan is to replicate their same processing, label them as is, then subset the treatment="RT"
############Seurat processing
############
#O_RT.data <- Read10X(data.dir = "~/SoupX/PATH/C57_RT_Rep1_LN_SoupX F:/new_analysis/integrated_analysis/bat/Burl2022/rep1/RT")
O_RT.data <- Read10X(data.dir = "F:/new_analysis/integrated_analysis/bat/Burl2022/linneg/rep1/RT")
O_RT <- CreateSeuratObject(counts = O_RT.data, min.cells=3, project = "C57_RT_Rep1_LN_SoupX")
O_RT$Replicate <- "Rep1"
O_RT$Treatment <- "CONTROL"
str(O_RT) #5886
############################################################################
#O_6C.data <- Read10X(data.dir = "~/SoupX/PATH/C57_COLD_Rep1_LN_SoupX F:/new_analysis/integrated_analysis/bat/Burl2022/rep1/COLD")
O_6C.data <- Read10X(data.dir = "F:/new_analysis/integrated_analysis/bat/Burl2022/linneg/rep1/COLD")
O_6C <- CreateSeuratObject(counts = O_6C.data, min.cells=3, project = "C57_COLD_Rep1_LN_SoupX")
O_6C$Replicate <- "Rep1"
O_6C$Treatment <- "COLD"
str(O_6C) #10242
############################################################################
#R_RT.data <- Read10X(data.dir = "~/SoupX/PATH/C57_RT_Rep2_LN_SoupX F:/new_analysis/integrated_analysis/bat/Burl2022/rep2/RT")
R_RT.data <- Read10X(data.dir = "F:/new_analysis/integrated_analysis/bat/Burl2022/linneg/rep2/RT")
R_RT <- CreateSeuratObject(counts = R_RT.data, min.cells=3, project = "C57_RT_Rep2_LN_SoupX")
R_RT$Replicate <- "Rep2"
R_RT$Treatment <- "CONTROL"
str(R_RT) #22072
############################################################################
#R_6C.data <- Read10X(data.dir = "~/SoupX/PATH/C57_COLD_Rep2_LN_SoupX F:/new_analysis/integrated_analysis/bat/Burl2022/rep2/COLD")
R_6C.data <- Read10X(data.dir = "F:/new_analysis/integrated_analysis/bat/Burl2022/linneg/rep2/COLD")
R_6C <- CreateSeuratObject(counts = R_6C.data, min.cells=3, project = "C57_COLD_Rep2_LN_SoupX")
R_6C$Replicate <- "Rep2"
R_6C$Treatment <- "COLD"
str(R_6C) #7414


###################
#Subset
###################
O_RT[["percent.mt"]] <- PercentageFeatureSet(O_RT, pattern = "^mt-")
O_6C[["percent.mt"]] <- PercentageFeatureSet(O_6C, pattern = "^mt-")
R_RT[["percent.mt"]] <- PercentageFeatureSet(R_RT, pattern = "^mt-")
R_6C[["percent.mt"]] <- PercentageFeatureSet(R_6C, pattern = "^mt-")
O_RT <- subset(O_RT, subset = percent.mt < 10 & nCount_RNA < 15000 & nFeature_RNA > 100)
O_6C <- subset(O_6C, subset = percent.mt < 10 & nCount_RNA < 15000 & nFeature_RNA > 100)
R_RT <- subset(R_RT, subset = percent.mt < 10 & nCount_RNA < 15000 & nFeature_RNA > 100)
R_6C <- subset(R_6C, subset = percent.mt < 10 & nCount_RNA < 15000 & nFeature_RNA > 100)
str(O_RT) #5603
str(O_6C) #9855
str(R_RT) #6171
str(R_6C) #7062


O_RT <- NormalizeData(O_RT)
O_6C <- NormalizeData(O_6C)
R_RT <- NormalizeData(R_RT)
R_6C <- NormalizeData(R_6C)


###################
#Merge
###################
#https://satijalab.org/seurat/v3.1/merge_vignette.html
OGnorm <- merge(O_RT, y = O_6C, project = "OG_Linneg", merge.data = TRUE)
RTnorm <- merge(R_RT, y = R_6C, project = "RT_Linneg", merge.data = TRUE)


###################
#Find Variable Features
###################
OGnorm <- FindVariableFeatures(OGnorm, selection.method = "vst", nfeatures = 2000)
RTnorm <- FindVariableFeatures(RTnorm, selection.method = "vst", nfeatures = 2000)

OGnorm <-SCTransform(OGnorm)
###################
###################
#Integrate
#######################dim = 12, res 0.1
LNanchors12 <- FindIntegrationAnchors(object.list = list(OGnorm, RTnorm), dims = 1:12) 
Linneg_mi_d12 <- IntegrateData(anchorset = LNanchors12, dims = 1:12) #this is not working with seurat V5. 

#need to use new layer function instead (https://github.com/satijalab/seurat/issues/8304) and
#https://github.com/satijalab/seurat/issues/7985#issuecomment-1806483174
OGnorm = JoinLayers(OGnorm)
LayerData(OGnorm, assay = "RNA", layer = "counts")

RTnorm= JoinLayers(RTnorm)
LayerData(RTnorm, assay = "RNA", layer = "counts")

#now do IntegrateData
LNanchors12 <- FindIntegrationAnchors(object.list = list(OGnorm, RTnorm), dims = 1:12) 
Linneg_mi_d12 <- IntegrateData(anchorset = LNanchors12, dims = 1:12) #this only integrate 2000 genes.

#I'll need to edit to integrate all genes instead
#3 options:
#1 adjust arg features.to.integrate= all genes in IntegrateData
#2 set VariableFeatures in FindIntegrationAnchors to be more than 2000
#3 merge control samples from very beginning, join layers, scale data


###################
#Scale
###################
allgenesmi12 <- rownames(Linneg_mi_d12)
Linneg_mi_d12 <- ScaleData(Linneg_mi_d12, features = allgenesmi12)
###################
#PCs
###################
Linneg_mi_d12 <- RunPCA(Linneg_mi_d12, features = VariableFeatures(object = Linneg_mi_d12))
Linneg_mi_d12 <- FindNeighbors(Linneg_mi_d12, dims = 1:12)
###################
#Visualize Data
###################
Linneg_mi_d12_r1 <- FindClusters(Linneg_mi_d12, resolution = 0.1)
#TSNE
Linneg_mi_d12_r1_t <- RunTSNE(Linneg_mi_d12_r1, dims = 1:12)
png("Linneg_mi_d12_r1_tsne_1.png")
DimPlot(Linneg_mi_d12_r1_t, reduction = "tsne")
dev.off()
png("Linneg_mi_d12_r1_tsne_2.png")
DimPlot(Linneg_mi_d12_r1_t, reduction = "tsne", split.by = "Treatment")
dev.off()


#change order of treatment for split.by
Linneg_mi_d12_r1_t$Treatment <- factor(Linneg_mi_d12_r1_t$Treatment, levels = c('CONTROL', 'COLD'))


#Set new cluster ids
#here I edit because the yellow ASC and red ASC are represented as 1 cluster in my processing.

new.cluster.ids <- c("ASC", "ASC", "VEC", "Prolif/Diff", "VSMC", "Immune", "Schwann")
names(new.cluster.ids) <- levels(Linneg_mi_d12_r1_t)
Linneg_mi_d12_r1_t <- RenameIdents(Linneg_mi_d12_r1_t, new.cluster.ids)
levels(Linneg_mi_d12_r1_t)
#[1] "ASC"         "VEC"         "Prolif/Diff" "VSMC"        "Immune"      "Schwann"    
Linneg_mi_d12_r1_t[["seurat_cident"]] <- Idents(object = Linneg_mi_d12_r1_t)



#create metadata file to subset to just ASCs
#I also edit here to fit my list to just get ASC and prolif/diff
Linneg_mi_d12_r1_t_ASCs <- subset(Linneg_mi_d12_r1_t, idents = c("ASC", "Prolif/Diff")) 
Linneg_mi_d12_r1_t@meta.data$seurat_cident %>% table()
#####paper ident
#ASC               1 ASC               2              ASC1-3                 VEC 
#               6665                5313                4801                4488 
#        Prolif/Diff                VSMC              Immune             Schwann 
#               2880                2561                1504                 479
####My ident
#ASC         VEC Prolif/Diff        VSMC      Immune     Schwann 
#16737        4404        2981        2563        1475         531 

Linneg_mi_d12_r1_t_ASCs@meta.data$seurat_cident %>% table()
#ASC               1 ASC               2              ASC1-3                 VEC 
#               6665                5313                4801                   0 
#        Prolif/Diff                VSMC              Immune             Schwann 
#               2880                   0                   0                   0

###My ident:
#ASC         VEC Prolif/Diff        VSMC      Immune     Schwann 
#16737           0        2981           0           0           0 
write.csv(Linneg_mi_d12_r1_t_ASCs@meta.data, "F:/new_analysis/integrated_analysis/bat/Burl2022/linneg/Linneg_mi_d12_r1_t_ASCs_md.csv")

##################################
##################################
##################################subset raw data to just ASCs; reanalyze
##################################
##################################
ad12 <- read.csv("F:/new_analysis/integrated_analysis/bat/Burl2022/linneg/Linneg_mi_d12_r1_t_ASCs_md.csv", header = TRUE)
str(ad12) #19659
          #ours is 19718
ad12_Rep1 <- ad12[ad12$Replicate == 'Rep1',]
ad12_Rep2 <- ad12[ad12$Replicate == 'Rep2',]
str(ad12_Rep1) #10766
              #ours is 10818

str(ad12_Rep2) #8893
              #ours is 8900

ad12_Rep1CON <- ad12_Rep1[ad12_Rep1$Treatment == 'CONTROL',]
str(ad12_Rep1CON) #3586
              #ours is 3596
head(ad12_Rep1CON) #_1_1
ad12_Rep1COLD <- ad12_Rep1[ad12_Rep1$Treatment == 'COLD',]
str(ad12_Rep1COLD) #7180
head(ad12_Rep1COLD) #_2_1

ad12_Rep2CON <- ad12_Rep2[ad12_Rep2$Treatment == 'CONTROL',]
str(ad12_Rep2CON) #3736
head(ad12_Rep2CON) #_1_2
ad12_Rep2COLD <- ad12_Rep2[ad12_Rep2$Treatment == 'COLD',]
str(ad12_Rep2COLD) #5157
head(ad12_Rep2COLD) #_2_2

ad12_Rep1CON$X <- gsub('_1_1', '', ad12_Rep1CON$X)
ad12_Rep1COLD$X <- gsub('_2_1', '', ad12_Rep1COLD$X)
ad12_Rep2CON$X <- gsub('_1_2', '', ad12_Rep2CON$X)
ad12_Rep2COLD$X <- gsub('_2_2', '', ad12_Rep2COLD$X)
head(ad12_Rep1CON)
head(ad12_Rep1COLD)
head(ad12_Rep2CON)
head(ad12_Rep2COLD)

######################
#subset to just ASCs
######################

O_RT_ad12_ASC <- subset(O_RT, cells = ad12_Rep1CON$X)
O_6C_ad12_ASC <- subset(O_6C, cells = ad12_Rep1COLD$X)
R_RT_ad12_ASC <- subset(R_RT, cells = ad12_Rep2CON$X)
R_6C_ad12_ASC <- subset(R_6C, cells = ad12_Rep2COLD$X)

str(O_RT_ad12_ASC) #3586
str(O_6C_ad12_ASC) #7180
str(R_RT_ad12_ASC) #3736
str(R_6C_ad12_ASC) #5157


###################
#Normalize
###################
O_RT_ad12_ASC <- NormalizeData(O_RT_ad12_ASC)
O_6C_ad12_ASC <- NormalizeData(O_6C_ad12_ASC)
R_RT_ad12_ASC <- NormalizeData(R_RT_ad12_ASC)
R_6C_ad12_ASC <- NormalizeData(R_6C_ad12_ASC)


###################
#Merge
###################
#https://satijalab.org/seurat/v3.1/merge_vignette.html
OGnorm_ad12_ASC <- merge(O_RT_ad12_ASC, y = O_6C_ad12_ASC, project = "OG_Linneg", merge.data = TRUE)
RTnorm_ad12_ASC <- merge(R_RT_ad12_ASC, y = R_6C_ad12_ASC, project = "RT_Linneg", merge.data = TRUE)


#join layers
#need to use new layer function instead (https://github.com/satijalab/seurat/issues/8304) and
#https://github.com/satijalab/seurat/issues/7985#issuecomment-1806483174
OGnorm_ad12_ASC = JoinLayers(OGnorm_ad12_ASC)
LayerData(OGnorm_ad12_ASC, assay = "RNA", layer = "counts")

RTnorm_ad12_ASC= JoinLayers(RTnorm_ad12_ASC)
LayerData(RTnorm_ad12_ASC, assay = "RNA", layer = "counts")

###################
#Find Variable Features
###################
OGnorm_ad12_ASC <- FindVariableFeatures(OGnorm_ad12_ASC, selection.method = "vst", nfeatures = 2000)
###
RTnorm_ad12_ASC <- FindVariableFeatures(RTnorm_ad12_ASC, selection.method = "vst", nfeatures = 2000)

#Integrate
###################
anchors8 <- FindIntegrationAnchors(object.list = list(OGnorm_ad12_ASC, RTnorm_ad12_ASC), dims = 1:8) 
ad12_d8 <- IntegrateData(anchorset = anchors8, dims = 1:8)
###################
#Scale
###################
ad12_genes8 <- rownames(ad12_d8)
ad12_d8 <- ScaleData(ad12_d8, features = ad12_genes8)
###################
#PCs
###################
ad12_d8 <- RunPCA(ad12_d8, features = VariableFeatures(object = ad12_d8))
ad12_d8 <- FindNeighbors(ad12_d8, dims = 1:8)
###################
#Visualize Data
###################
ad12_d8_r2 <- FindClusters(ad12_d8, resolution = 0.2)
#TSNE
ad12_d8_r2_t <- RunTSNE(ad12_d8_r2, dims = 1:8)
png("ad12_d8_r2_tsne_1.png")
DimPlot(ad12_d8_r2_t, reduction = "tsne")
dev.off()
DimPlot(ad12_d8_r2_t, split.by='Treatment', group.by="Replicate")

###################
#Set new cluster ids
###################
new.cluster.ids <- c("ASC1_cold-activated", "ASC2", "ASC1_quiescent", "ASC3", "Prolif/Diff",  "Prolif/Non-diff")
names(new.cluster.ids) <- levels(ad12_d8_r2_t)
ad12_d8_r2_t <- RenameIdents(ad12_d8_r2_t, new.cluster.ids)
levels(ad12_d8_r2_t)
#[1] "ASC1_cold-activated" "ASC2"                "ASC1_quiescent"     
#[4] "Prolif/Diff"         "ASC3"                "Prolif/Non-diff"
levels(x = ad12_d8_r2_t) <- c("ASC1_quiescent", "ASC1_cold-activated", "ASC2", "ASC3", "Prolif/Non-diff", "Prolif/Diff")
levels(ad12_d8_r2_t)

ad12_d8_r2_t@meta.data$burl_clusters <- as.factor(ad12_d8_r2_t@active.ident)
saveRDS(ad12_d8_r2_t, file="F:/new_analysis/integrated_analysis/bat/Burl2022/linneg/ASC_labeled.rds")

```




# Part B. Perform reference mapping

###################
#Reference mapping#
###################
## Load in BAT
```{r}
bat_cc <- readRDS('PATH/TO/bat_cc.rds') #this paper's bat_cc object
#if object already loaded, just change their name
bat<- bat_cc
```

## renames objects from above
```{r}
bat <- bat_cc
karlina <- karlina.preads
burl <- ad12_d8_r2_t
```

## Figure 5A BAT reference mapping, along with their counts for Fig. 5D

### Reference mapping for sarvari dataset
```{r}
bat_cc <- RunUMAP(bat_cc, dims=1:10, return.model = TRUE) #have to return the model to 
bat=bat_cc

DefaultAssay(bat) <- 'RNA'


burl_anchors <- FindTransferAnchors(
  reference = bat,
  query = burl,
  normalization.method = "LogNormalize",
  reference.reduction = "pca",
  dims = 1:50
) #1199 anchors

bat_ref <- MapQuery(
  anchorset = burl_anchors,
  query = burl,
  reference = bat,
  refdata = list(
    bat_bclusters= 'bat_named_bclusters'
  ),
  reference.reduction = "pca", 
  reduction.model = "umap"
)

DimPlot(bat_ref, reduction='ref.umap', group.by= 'predicted.bat_bclusters')




#counting
results <- data.frame()  # Initialize an empty data frame to store results

for (cluster in unique(bat_ref$predicted.bat_bclusters)) {
  # Get counts for each category within the current cluster
  cellcounts <- table(bat_ref$burl_clusters[bat_ref$predicted.bat_bclusters == cluster])
  
  # Combine the counts into a temporary data frame
  temp_df <- data.frame(
    predicted_cell_type = cluster,
    query_cells = names(cellcounts),
    count = as.integer(cellcounts)
  )
  
  # Append to the results data frame
  results <- rbind(results, temp_df)
}

#Fig. 5D Burl predicted labels counting
cell_counts <- table(bat_ref$burl_clusters, bat_ref$predicted.bat_bclusters)

write.csv(results, "R:/Galmozzi lab/Hoang/Projects/Adipocyte Subtypes/integrated_analysis/Reference_mapping/sarvari_ref_counts.csv")
write.csv(results, "F:/new_analysis/labeled_analysis/reference_mapping_rds/burl_ref_counts.csv") #counts
write.csv(cell_counts, "F:/new_analysis/labeled_analysis/reference_mapping_rds/burl_ref_counts2.csv")

write.csv(results, "R:/Galmozzi lab/Hoang/Projects/Adipocyte Subtypes/integrated_analysis/Reference_mapping/burl_ref_counts.csv")

#merge reference and query
reference=bat
query=bat_ref
reference$id <- 'reference'
query$id <- 'query'
refquery <- merge(reference, query)
refquery[["umap"]] <- merge(reference[["umap"]], query[["ref.umap"]])

DimPlot(refquery, group.by = 'id', shuffle = TRUE)

burl_colors= c("red","darkgoldenrod4","brown","violet","blue","springgreen4")

DimPlot(refquery, group.by='burl_clusters', pt.size = 3, cols=burl_colors, na.value='lightgrey')

DimPlot(refquery, group.by='predicted.bat_bclusters')



#setting the right order
refquery <-SetIdent(refquery, value='predicted.bat_bclusters')
my_levels<- c('Early Preads B1', "Early Preads B2", "Early Preads B3","Early Preads B5", "bCAPs",'bFAPLs')
Idents(refquery) <- factor(Idents(refquery), levels=my_levels)

# Fig. 5C Burl reference mapping

DimPlot(refquery, pt.size=3, cols = c("#00868b", "#cd5b45", "#FFE300", "#00FFFF", "#0000FF", "#FF0000")) #with ordering
DimPlot(refquery, pt.size=3, cols = c( "#0000FF", "#FF0000", "#00868b", "#cd5b45", "#FFE300", "#00FFFF"), group.by="predicted.bat_bclusters", na.value='lightgrey')


saveRDS(bat_ref, file= 'F:/new_analysis/labeled_analysis/reference_mapping_rds/mapped_burl.rds')
saveRDS(refquery, file= 'F:/new_analysis/labeled_analysis/reference_mapping_rds/mapped_burl_merged_umap.rds')

```


### Reference mapping for Karlina dataset
```{r}
levels(karlina)
new_cluster_ids <- c('0(Karlina 0)', '1(Karlina 3)', "2(Karlina 2)", '3(Karlina 1)')

names(new_cluster_ids) <- levels(karlina)
karlina <- RenameIdents(karlina, new_cluster_ids)
karlina@meta.data$karlina_reconstructed_clusters <- as.factor(karlina@active.ident)

DefaultAssay(bat) <- "SCT"

karlina_anchors <- FindTransferAnchors(
  reference = bat,
  query = karlina,
  normalization.method = "LogNormalize",
  reference.reduction = "pca",
  dims = 1:50
) #265 anchors

bat_ref <- MapQuery(
  anchorset = karlina_anchors,
  query = karlina,
  reference = bat,
  refdata = list(
    bat_bclusters= 'bat_named_bclusters'
  ),
  reference.reduction = "pca", 
  reduction.model = "umap"
)

DimPlot(bat_ref, reduction='ref.umap', group.by= 'predicted.bat_bclusters')

#plotting on the reference UMAP
reference= bat
reference <- DietSeurat(reference, counts = FALSE, dimreducs = c("umap"))

query <- DietSeurat(bat_ref, counts = FALSE, dimreducs = c("ref.umap"))

results <- data.frame()  # Initialize an empty data frame to store results

for (cluster in unique(bat_ref$predicted.bat_bclusters)) {
  # Get counts for each category within the current cluster
  cellcounts <- table(bat_ref$karlina_reconstructed_clusters[bat_ref$predicted.bat_bclusters == cluster])
  
  # Combine the counts into a temporary data frame
  temp_df <- data.frame(
    predicted_cell_type = cluster,
    query_cells = names(cellcounts),
    count = as.integer(cellcounts)
  )
  
  # Append to the results data frame
  results <- rbind(results, temp_df)
}

#Fig. 5D Karlina
cell_counts <- table(bat_ref$karlina_reconstructed_clusters, bat_ref$predicted.bat_bclusters)
write.csv(results, "R:/Galmozzi lab/Hoang/Projects/Adipocyte Subtypes/integrated_analysis/Reference_mapping/karlina_ref_counts.csv")

write.csv(results, "F:/new_analysis/labeled_analysis/reference_mapping_rds/karlina_ref_counts.csv") #counts
write.csv(cell_counts, "F:/new_analysis/labeled_analysis/reference_mapping_rds/karlina_ref_counts2.csv")

#merge reference and query
reference$id <- 'reference'
query$id <- 'query'
refquery <- merge(reference, query)
refquery[["umap"]] <- merge(reference[["umap"]], query[["ref.umap"]])
#refquery <- RunUMAP(refquery, reduction = 'pca', dims = 1:50)


#Fig. 5C Karlina
DimPlot(refquery, group.by = 'id', shuffle = TRUE)
karlina_colors= c('aquamarine4','plum4','navy','violetred')
DimPlot(refquery, group.by='karlina_reconstructed_clusters', pt.size = 3, cols=karlina_colors, na.value='lightgrey')

DimPlot(refquery, group.by='predicted.bat_bclusters', pt.size=3, cols = c('#FF0000','#cd5b45'), na.value = 'lightgrey')

saveRDS(refquery, file= 'F:/new_analysis/labeled_analysis/reference_mapping_rds/mapped_karlina_merged_umap.rds')
saveRDS(bat_ref, file= 'F:/new_analysis/labeled_analysis/reference_mapping_rds/mapped_karlina.rds')
